---
title: "CLOVERS_Paper"
author: "Jordan Lo"
date: "2023-09-08"
output:
  word_document: default
  html_document: default
---

```{r, include = FALSE, echo = FALSE}
knitr::opts_chunk$set(echo = FALSE, include = FALSE, warning = FALSE)
```

```{r}
# tidyverse allows us to do our data processing
library("tidyverse")
# readxl allows us to read excel files
library("readxl")
# writexl allows us to write to excel files
library("writexl")
# qpcR lets us merge data frames side by side
library("qpcR")
# car gives us access to more variance tests
library("car")
# ggpubr adds ways to customize ggplots
library("ggpubr")
# sandwich lets us find a robust std error 
library("sandwich")
# parameters extracts model parameters
library("parameters")
```

```{r}
chr_urn_old <- read_excel("B:\\CHROME\\Urine_Biomarker_V1\\Ang-2_sFAS_2022-04\\Ang-2_Data\\CHR_US_V1_Ang-2_MSD_Plates1-6_Data_NO_SKS_JH_4-7-2022.xlsx",
                          sheet = "RawData")
colnames(chr_urn_old)[15] <- "Calc. Concentration"
chr_urn_old$Sample <- as.numeric(chr_urn_old$Sample)
chr_urn_old$`Calc. Concentration` <- as.numeric(chr_urn_old$`Calc. Concentration`)

chr_urn_new <- read_excel("B:\\CHROME\\Urine_Biomarker_V1\\Ang-2_sFAS_2022-04\\Ang-2_Data\\CHR_US_V1_Ang2_Plate1 (redo 12.7.22).xlsx",
                          sheet = "RawData")
```

```{r}
# plots the plate from the old CHROME run to find the plate that was higher than all the others
plot <- ggplot(data = chr_urn_old,
               mapping = aes(x = `Plate Name`,
                             y = log(`Calc. Concentration`))) +
  geom_boxplot()
plot

# filters out said plate
chr_urn_old <- chr_urn_old %>%
  filter(`Plate Name` != "Plate1b_23L0EAV560")
```

```{r}
# merges the old chr data with the re-run plate 1

chr_data <- merge(chr_urn_old,
                  chr_urn_new,
                  all.x = TRUE,
                  all.y = TRUE)

# subsets the dataframe into the required columns
chr_data_conc <- chr_data[,c("Plate Name",
                             "Sample", 
                             "Calc. Conc. Mean")]

# converts columns to numeric and drops the resultant NA values
chr_data_conc$Sample <- as.numeric(chr_data_conc$Sample)
#sum(is.na(chr_data_conc$Sample))
chr_data_conc$`Calc. Conc. Mean` <- as.numeric(chr_data_conc$`Calc. Conc. Mean`)
colnames(chr_data_conc) <- c("plate_name",
                             "study_id",
                             "ang2_concentration")

# removes duplicate samples in the data
chr_data_conc <- chr_data_conc %>%
  distinct(study_id,
           .keep_all = TRUE)

# imputes a value for values below LLOD (1/2 LLOD of Ang-2 Assay)
chr_data_conc$ang2_concentration[is.nan(chr_data_conc$ang2_concentration)] <- 5

# removes remaining NA values
chr_data_conc <- na.omit(chr_data_conc)

# reads in Leila's provided clinical data set
clinical <- read_csv("chrome_clinical_aki_data_for_jordan_zoie_2023_09_05.csv")

# creates a separate dataframe of the CHROME/CITRC data and recodes the study_ID to the appropriate CHROME ID
chr_conc_citrc  <- chr_data_conc %>%
  filter(study_id > 522)
colnames(chr_conc_citrc)[2] <- "citrc_studyid" 
chr_conc_citrc <- merge(chr_conc_citrc,
                        clinical)
chr_conc_citrc <- chr_conc_citrc[, c(2,4,3)]

# combines the CHROME/CITRC data with the original biomarker dataframe
chr_conc_clinical <- rbind(chr_conc_citrc,
                           chr_data_conc)

# merges the clinical and biomarker dataframes
chr_conc_clinical <- merge(chr_conc_clinical,
                           clinical)
```

<br>
There are 201 patients that remain after merging the two datasets and filtering out those with clinical data and Ang-2 biomarker data available

```{r}
# splits data into tertiles based on ang-2 concentrations and repositions the column
chr_conc_clinical$ang2_tertile <- as.numeric(cut_number(chr_conc_clinical$ang2_concentration,
                                        3))
chr_conc_clinical <- chr_conc_clinical %>%
  relocate(244, 
           .after = 3)
#chr_conc_clinical %>% count(ang2_tertile)

# creates a log2 column
chr_conc_clinical <- chr_conc_clinical %>%
                     mutate(ang2_concentration_log2 = log2(ang2_concentration)) %>%
  relocate(245, 
           .after = 3)
```

<br>

```{r, include = TRUE}
# creates a histogram of the ang-2 concentrations
histogram <- ggplot(data = chr_conc_clinical,
                    mapping = aes(x = ang2_concentration)) +
  geom_histogram()
histogram

# colnames(chr_conc_clinical)

chr_conc_clinical$aki_hosp <- as.character(chr_conc_clinical$aki_hosp)
# boxplot of ang-2 concentrations and AKI within 7 days
ang_7d_aki <- ggplot(data = chr_conc_clinical,
                     mapping = aes(x = aki_hosp,
                                   y = ang2_concentration_log2)) + 
  geom_boxplot()
ang_7d_aki

chr_conc_clinical$dialysis_study <- as.character(chr_conc_clinical$dialysis_study)
# boxplot of ang-2 conc and dialysis within 7 days
ang_7d_dialysis <- ggplot(data = chr_conc_clinical,
                          mapping = aes(x = dialysis_study,
                                        y = ang2_concentration_log2)) +
  geom_boxplot()
ang_7d_dialysis
```

```{r}
# normalizing ang-2 to urine creatinine levels

# reads in lab data for CHROME
lab <- read_csv(file = "chrome_lab_data_for_jordan_zoie_2023_09_12.csv")
lab <- lab %>%
  filter(Visit == "V1") %>%
  dplyr::select(study_id,
         ucr_mg_dL,
         ukim1_pg_mL)

# merges lab data and clinical data
clinical_lab_msd <- merge(chr_conc_clinical,
                          lab)

# creates a new variable of the ang-2 concentration normalized to urine creatinine (pg/mg)
clinical_lab_msd <- clinical_lab_msd %>%
  mutate(ucr_norm_ang2 = (ang2_concentration/(ucr_mg_dL*(1/100))))
```

```{r}
# creates a histogram of the normalized ang-2 concentrations
clinical_lab_msd_no_outlier <- clinical_lab_msd %>%
  filter(study_id != 278)
  
histogram <- ggplot(data = clinical_lab_msd,
                    mapping = aes(x = ucr_norm_ang2)) +
  geom_histogram()
histogram

histogram <- ggplot(data = clinical_lab_msd_no_outlier,
                    mapping = aes(x = ucr_norm_ang2)) +
  geom_histogram()
histogram

clinical_lab_msd$aki_hosp <- as.character(clinical_lab_msd$aki_hosp)
# boxplot of normalized ang-2 concentrations and AKI within 7 days
ang_hosp_aki <- ggplot(data = clinical_lab_msd,
                     mapping = aes(x = aki_hosp,
                                   y = log2(ucr_norm_ang2))) + 
  geom_boxplot()
ang_hosp_aki

clinical_lab_msd$dialysis_study <- as.character(clinical_lab_msd$dialysis_study)
# boxplot of normalized ang-2 conc and dialysis within 7 days
ang_7d_dialysis <- ggplot(data = clinical_lab_msd,
                          mapping = aes(x = dialysis_study,
                                        y = log2(ucr_norm_ang2))) +
  geom_boxplot()
ang_7d_dialysis
```

```{r, include = TRUE}
# 9/27/23

# Correlation plot of ucr_norm_ang2 to Kim1 for Zoie to look at
norm_ang2_kim1 <- ggplot(data = clinical_lab_msd,
                         mapping = aes(
                           y = log2(ukim1_pg_mL),
                           x = log2(ucr_norm_ang2)))

norm_ang2_kim1 + 
  geom_smooth(method = "lm",
              se = FALSE) + 
  geom_point() +
  stat_regline_equation() +
  stat_cor(label.y = 15,
           aes(label = ..r.label..))
```
Looking into the data by tertiles based on Ang-2 normalized to urine creatinine levels
```{r}
# splitting data into tertiles on ucr_norm_ang2
clinical_lab_msd$ucr_norm_ang2_tertile <- as.numeric(cut_number(clinical_lab_msd$ucr_norm_ang2,
                                        3))
#clinical_lab_msd %>% count(ucr_norm_ang2_tertile)
clinical_lab_msd$ucr_norm_ang2_tertile <- as.factor(clinical_lab_msd$ucr_norm_ang2_tertile)
clinical_lab_msd$aki_hosp <- as.factor(clinical_lab_msd$aki_hosp)
```

```{r, include = TRUE}
# Looks at the 7 day aki split by normalized ang-2 tertiles
aki_norm_ang2 <- ggplot(data = clinical_lab_msd,
                        mapping = aes(fill = ucr_norm_ang2_tertile,
                                      x = aki_hosp))
aki_norm_ang2 + 
  geom_bar(position = "dodge")

# Looks at the 7 day dialysis split by normalized ang-2 tertiles
dialysis_norm_ang2 <- ggplot(data = clinical_lab_msd,
                             mapping = aes(fill = ucr_norm_ang2_tertile,
                                           x = dialysis_study))
dialysis_norm_ang2 +
  geom_bar(position = "dodge")

boxes_norm_ang2 <- ggplot(data = clinical_lab_msd,
                          mapping = aes(x = ucr_norm_ang2_tertile,
                                        y = log2(ucr_norm_ang2)))
boxes_norm_ang2 + geom_boxplot()
```

```{r, include = TRUE}
# 10/5/2023

# reads in the urine ngal msd data
ngal <- read_excel("B:/CHROME/Urine_Biomarker_V1/IL-18_NGAL_2022-03/NGAL_Data/CHR_US_V1_NGAL_MSD_Plates1-6_Data_NO__SKS_3-16-2022.xlsx",
                   sheet = "WideSampleData") %>%
  dplyr::select(c("Sample",
                  "NGAL (pg/mL)"))

# renames columns in ngal dataset and merges with the clinical_lab_msd data
colnames(ngal) <- c("study_id",
                    "ungal_concentration")
clinical_lab_msd <- left_join(clinical_lab_msd,
                               ngal,
                               by = "study_id",
                               all.x = TRUE)

# generates a scatterplot of ngal vs normalized ang2
ngal_uang2 <- ggplot(data = clinical_lab_msd,
                     mapping = aes(x = log2(ucr_norm_ang2),
                                   y = log2(ungal_concentration)))

ngal_uang2  + 
  geom_smooth(method = "lm",
              se = FALSE) + 
  geom_point() +
  stat_regline_equation(label.y = 23) +
  stat_cor(label.y = 22,
           aes(label = ..r.label..))
```

```{r, include = TRUE}
pang2 <- read_excel("B:/CHROME/Plasma/Ang1Ang2/2023-02-08/CHR_Ang1Ang2_20230209_ALL_PLATES_ANALYZED.xlsx",
                    sheet = "WideSampleData")
colnames(pang2) <- c("study_id",
                     "PE_ang1_v1_concentration",
                     "PE_ang2_v1_concentration")

clinical_lab_msd <- left_join(clinical_lab_msd,
                               pang2,
                               by = "study_id",
                               all.x = TRUE)

clinical_lab_msd$PE_ang1_v1_concentration <- as.numeric(clinical_lab_msd$PE_ang1_v1_concentration)
clinical_lab_msd$PE_ang2_v1_concentration <- as.numeric(clinical_lab_msd$PE_ang2_v1_concentration)


# generates a scatterplot of ngal vs normalized ang2
peang2_uang2 <- ggplot(data = clinical_lab_msd,
                     mapping = aes(x = log2(ucr_norm_ang2),
                                   y = log2(PE_ang2_v1_concentration)))

peang2_uang2  + 
  geom_smooth(method = "lm",
              se = FALSE) + 
  geom_point() +
  stat_regline_equation(label.y = 15.5) +
  stat_cor(label.y = 15,
           aes(label = ..r.label..))
```

```{r, include = TRUE, echo = TRUE}
# 10/6/23

# finds overlap between ang2 and normalized ang 2 tertiles
ang2_tertile <- clinical_lab_msd %>%
  dplyr::select(c("study_id",
                  "ang2_tertile")) %>%
  filter(ang2_tertile == 3)

norm_ang2_tertile <- clinical_lab_msd %>%
  dplyr::select(c("study_id",
                  "ucr_norm_ang2_tertile")) %>%
  filter(ucr_norm_ang2_tertile == 3)

tertile_overlap <- inner_join(ang2_tertile,
                              norm_ang2_tertile,
                              by = "study_id")

length(tertile_overlap$study_id)
```
<br>
We see very high overlap between the highest Ang-2 Concentration tertiles and the highest normalized Ang-2 tertiles. 
53/64 patients are in the highest tertiles for both.
<br>

```{r}
# # table 1 data finding
# colnames(clinical_lab_msd)
# 
# # custom function to find demographics broken down by tertile
# table_1_count <- function(tertile, column) {
#   clinical_lab_msd %>%
#     filter(ucr_norm_ang2_tertile == tertile) %>%
#     count(column)
# }
# 
# tertile <- c(1,2,3)
# 
# # finds mean age by tertile
# clinical_lab_msd %>%
#   group_by(ucr_norm_ang2_tertile) %>%
#   summarise_at(vars(age),
#                list(name = mean))
# 
# # finds number of males in each tertile
# lapply(tertile,
#        table_1_count,
#        column = "male")
# 
# # finds racial demographics in each tertile
# lapply(tertile,
#        table_1_count,
#        column = "race")
# 
# # finds tertile breakdown of hosp_aki
# lapply(tertile,
#        table_1_count,
#        column = "aki_hosp_stage")
# 
# # tertile breakdown of 7d_dialysis
# lapply(tertile,
#        table_1_count,
#        column = "dialysis_study")
# 
# # tertile breakdown for ckd
# lapply(tertile,
#        table_1_count,
#        column = "cri")
# 
# clinical_lab_msd %>%
#   summarise_at(vars(age),
#                list(name = sd))
```

```{r}
# odds ratio
# splits out data into individual tertiles
or_lt <- clinical_lab_msd %>%
  filter(ucr_norm_ang2_tertile == 1)
or_mt <- clinical_lab_msd %>%
  filter(ucr_norm_ang2_tertile == 2)
or_ht <- clinical_lab_msd %>%
  filter(ucr_norm_ang2_tertile == 3)

# new variable establishing lowest tertile as a reference 
or_lt$tertile_or <- 0
or_mt$tertile_or <- 1
or_ht$tertile_or <- 1

# binds together the tertiles for analysis
or_mt_lt <- rbind(or_lt,
                 or_mt)
or_ht_lt <- rbind(or_lt,
                  or_ht)
or_mt_ht <- rbind(or_mt,
                  or_ht)

# runs a binomial regression with several variables for covariate adjustment 
result <- glm(aki_hosp ~ tertile_or + age + male, 
              data = or_mt_lt,
              family = "binomial")

summary(result)
exp(cbind(coef(result), confint.default(result)))

result <- glm(aki_hosp ~ tertile_or + age + male, 
              data = or_ht_lt,
              family = "binomial")

summary(result) 
exp(cbind(coef(result), confint.default(result)))
```


```{r}
# relative risk

# converts several variables from factor to numeric
clinical_lab_msd$aki_hosp <- as.numeric(clinical_lab_msd$aki_hosp)
clinical_lab_msd$dialysis_study <- as.numeric(clinical_lab_msd$dialysis_study)

# splits out data into individual tertiles
rr_lt <- clinical_lab_msd %>%
  filter(ucr_norm_ang2_tertile == 1)
rr_mt <- clinical_lab_msd %>%
  filter(ucr_norm_ang2_tertile == 2)
rr_ht <- clinical_lab_msd %>%
  filter(ucr_norm_ang2_tertile == 3)

# new variable establishing lowest tertile as a reference for relative risk
rr_lt$tertile_rr <- 0
rr_mt$tertile_rr <- 1
rr_ht$tertile_rr <- 1
rr_lt$tertile_rr_nb <- 0
rr_mt$tertile_rr_nb <- 0
rr_ht$tertile_rr_nb <- 1

# binds together the tertiles for analysis
rr_mt_lt <- rbind(rr_lt,
                 rr_mt)
rr_ht_lt <- rbind(rr_lt,
                  rr_ht)
rr_mt_ht <- rbind(rr_mt,
                  rr_ht)
```


```{r, include = TRUE, echo = TRUE}
# runs a poisson regression with covariate adjustment to find relative risk

# aki_hosp
# regression on lowest/middle tertile with adjustment
result <- glm(aki_hosp ~ tertile_rr + age + male + bmi + mv + covid_confirmed, 
              data = rr_mt_lt,
              family = "poisson")

model_parameters(result,
                 exponentiate = TRUE,
                 vcov = "HC",
                 vcov_args = list(type = "HC0"))

# unadjusted model
result <- glm(aki_hosp ~ tertile_rr, 
              data = rr_mt_lt,
              family = "poisson")

model_parameters(result,
                 exponentiate = TRUE,
                 vcov = "HC",
                 vcov_args = list(type = "HC0"))

# regression on lowest/highest tertile with adjustment 
result <- glm(aki_hosp ~ tertile_rr + age + male + bmi + mv + covid_confirmed, 
              data = rr_ht_lt,
              family = "poisson")

model_parameters(result,
                 exponentiate = TRUE,
                 vcov = "HC",
                 vcov_args = list(type = "HC0"))

# unadjusted 
result <- glm(aki_hosp ~ tertile_rr, 
              data = rr_ht_lt,
              family = "poisson")

model_parameters(result,
                 exponentiate = TRUE,
                 vcov = "HC",
                 vcov_args = list(type = "HC0"))

```
We do see that even without adjustments, being in the top tertile of the dataset predicts higher chances of developing aki in the hospital.

```{r, include = TRUE, echo = TRUE}
# dialysis_study
# regression on lowest/middle tertile with adjustment
result <- glm(dialysis_study ~ tertile_rr + age + male + bmi + mv + covid_confirmed, 
              data = rr_mt_lt,
              family = "poisson")

model_parameters(result,
                 exponentiate = TRUE,
                 vcov = "HC",
                 vcov_args = list(type = "HC0"))

# unadjusted
result <- glm(dialysis_study ~ tertile_rr, 
              data = rr_mt_lt,
              family = "poisson")

model_parameters(result,
                 exponentiate = TRUE,
                 vcov = "HC",
                 vcov_args = list(type = "HC0"))


# regression on lowest/highest tertile with adjustment
result <- glm(dialysis_study ~ tertile_rr + age + male + bmi + mv + covid_confirmed, 
              data = rr_ht_lt,
              family = "poisson")

model_parameters(result,
                 exponentiate = TRUE,
                 vcov = "HC",
                 vcov_args = list(type = "HC0"))

# unadjusted
result <- glm(dialysis_study ~ tertile_rr, 
              data = rr_ht_lt,
              family = "poisson")

model_parameters(result,
                 exponentiate = TRUE,
                 vcov = "HC",
                 vcov_args = list(type = "HC0"))
```

```{r}
# death_hospital
# regression on lowest/middle tertile with adjustment
result <- glm(death_hospital ~ tertile_rr + age + male + bmi + mv + covid_confirmed, 
              data = rr_mt_lt,
              family = "poisson")

model_parameters(result,
                 exponentiate = TRUE,
                 vcov = "HC",
                 vcov_args = list(type = "HC0"))

# unadjusted
result <- glm(death_hospital ~ tertile_rr, 
              data = rr_mt_lt,
              family = "poisson")

model_parameters(result,
                 exponentiate = TRUE,
                 vcov = "HC",
                 vcov_args = list(type = "HC0"))


# regression on lowest/highest tertile with adjustment
result <- glm(death_hospital ~ tertile_rr + age + male + bmi + mv + covid_confirmed, 
              data = rr_ht_lt,
              family = "poisson")

model_parameters(result,
                 exponentiate = TRUE,
                 vcov = "HC",
                 vcov_args = list(type = "HC0"))

# unadjusted
result <- glm(death_hospital ~ tertile_rr, 
              data = rr_ht_lt,
              family = "poisson")

model_parameters(result,
                 exponentiate = TRUE,
                 vcov = "HC",
                 vcov_args = list(type = "HC0"))
```

