---
title: "CLOVERS_Paper"
author: "Jordan Lo"
date: "2023-09-08"
output:
  word_document: default
  html_document: default
---

```{r, include = FALSE, echo = FALSE}
knitr::opts_chunk$set(echo = FALSE, include = FALSE, warning = FALSE)
```

```{r}
# tidyverse allows us to do our data processing
library("tidyverse")
# readxl allows us to read excel files
library("readxl")
# writexl allows us to write to excel files
library("writexl")
# qpcR lets us merge data frames side by side
library("qpcR")
# car gives us access to more variance tests
library("car")
# ggpubr adds ways to customize ggplots
library("ggpubr")
# sandwich lets us find a robust std error 
library("sandwich")
# parameters extracts model parameters
library("parameters")
```

```{r}
chr_urn_old <- read_excel("B:\\CHROME\\Urine_Biomarker_V1\\Ang-2_sFAS_2022-04\\Ang-2_Data\\CHR_US_V1_Ang-2_MSD_Plates1-6_Data_NO_SKS_JH_4-7-2022.xlsx",
                          sheet = "RawData")
colnames(chr_urn_old)[15] <- "Calc. Concentration"
chr_urn_old$Sample <- as.numeric(chr_urn_old$Sample)
chr_urn_old$`Calc. Concentration` <- as.numeric(chr_urn_old$`Calc. Concentration`)

chr_urn_new <- read_excel("B:\\CHROME\\Urine_Biomarker_V1\\Ang-2_sFAS_2022-04\\Ang-2_Data\\CHR_US_V1_Ang2_Plate1 (redo 12.7.22).xlsx",
                          sheet = "RawData")
```

```{r}
# plots the plate from the old CHROME run to find the plate that was higher than all the others
plot <- ggplot(data = chr_urn_old,
               mapping = aes(x = `Plate Name`,
                             y = log(`Calc. Concentration`))) +
  geom_boxplot()
plot

# filters out said plate
chr_urn_old <- chr_urn_old %>%
  filter(`Plate Name` != "Plate1b_23L0EAV560")
```

```{r}
# merges the old chr data with the re-run plate 1

chr_data <- merge(chr_urn_old,
                  chr_urn_new,
                  all.x = TRUE,
                  all.y = TRUE)

# subsets the dataframe into the required columns
chr_data_conc <- chr_data[,c("Plate Name",
                             "Sample", 
                             "Calc. Conc. Mean")]

# converts columns to numeric and drops the resultant NA values
chr_data_conc$Sample <- as.numeric(chr_data_conc$Sample)
#sum(is.na(chr_data_conc$Sample))
chr_data_conc$`Calc. Conc. Mean` <- as.numeric(chr_data_conc$`Calc. Conc. Mean`)
colnames(chr_data_conc) <- c("plate_name",
                             "study_id",
                             "ang2_concentration")

# removes duplicate samples in the data
chr_data_conc <- chr_data_conc %>%
  distinct(study_id,
           .keep_all = TRUE)

# imputes a value for values below LLOD (1/2 LLOD of Ang-2 Assay)
chr_data_conc$ang2_concentration[is.nan(chr_data_conc$ang2_concentration)] <- 5

# removes remaining NA values
chr_data_conc <- na.omit(chr_data_conc)

# reads in Leila's provided clinical data set
clinical <- read_csv("chrome_clinical_aki_data_for_jordan_zoie_2023_09_05.csv")

# creates a separate dataframe of the CHROME/CITRC data and recodes the study_ID to the appropriate CHROME ID
chr_conc_citrc  <- chr_data_conc %>%
  filter(study_id > 522)
colnames(chr_conc_citrc)[2] <- "citrc_studyid" 
chr_conc_citrc <- merge(chr_conc_citrc,
                        clinical)
chr_conc_citrc <- chr_conc_citrc[, c(2,4,3)]

# combines the CHROME/CITRC data with the original biomarker dataframe
chr_conc_clinical <- rbind(chr_conc_citrc,
                           chr_data_conc)

# merges the clinical and biomarker dataframes
chr_conc_clinical <- merge(chr_conc_clinical,
                           clinical)
```

<br>
There are 201 patients that remain after merging the two datasets and filtering out those with clinical data and Ang-2 biomarker data available

```{r}
# creates a log2 column
chr_conc_clinical <- chr_conc_clinical %>%
                     mutate(ang2_concentration_log2 = log2(ang2_concentration)) %>%
  relocate(244, 
           .after = 3)
```

<br>

```{r, include = TRUE}
# creates a histogram of the ang-2 concentrations
histogram <- ggplot(data = chr_conc_clinical,
                    mapping = aes(x = ang2_concentration)) +
  geom_histogram()
histogram

# colnames(chr_conc_clinical)

chr_conc_clinical$aki_hosp <- as.character(chr_conc_clinical$aki_hosp)
# boxplot of ang-2 concentrations and AKI within 7 days
ang_7d_aki <- ggplot(data = chr_conc_clinical,
                     mapping = aes(x = aki_hosp,
                                   y = ang2_concentration_log2)) + 
  geom_boxplot()
ang_7d_aki

chr_conc_clinical$dialysis_study <- as.character(chr_conc_clinical$dialysis_study)
# boxplot of ang-2 conc and dialysis within 7 days
ang_7d_dialysis <- ggplot(data = chr_conc_clinical,
                          mapping = aes(x = dialysis_study,
                                        y = ang2_concentration_log2)) +
  geom_boxplot()
ang_7d_dialysis
```

```{r}
# normalizing biomarkers to urine creatinine levels

# reads in lab data for CHROME
lab <- read_csv(file = "chrome_lab_data_for_jordan_zoie_2023_09_12.csv")
lab <- lab %>%
  filter(Visit == "V1") %>%
  dplyr::select(study_id,
         ucr_mg_dL,
         ukim1_pg_mL)

# merges lab data and clinical data
clinical_lab_msd <- merge(chr_conc_clinical,
                          lab)

# colnames(clinical_lab_msd)

# splits data into tertiles based on ang-2 concentrations and repositions the column
clinical_lab_msd$ang2_tertile <- as.numeric(cut_number(clinical_lab_msd$ang2_concentration,
                                        3))
clinical_lab_msd <- clinical_lab_msd %>%
  relocate(244, 
           .after = 3)
clinical_lab_msd %>% count(ang2_tertile)

# reads in the urine ngal msd data
ngal <- read_excel("B:/CHROME/Urine_Biomarker_V1/IL-18_NGAL_2022-03/NGAL_Data/CHR_US_V1_NGAL_MSD_Plates1-6_Data_NO__SKS_3-16-2022.xlsx",
                   sheet = "WideSampleData") %>%
  dplyr::select(c("Sample",
                  "NGAL (pg/mL)"))

# renames columns in ngal dataset and merges with the clinical_lab_msd data
colnames(ngal) <- c("study_id",
                    "ungal_concentration")
clinical_lab_msd <- left_join(clinical_lab_msd,
                               ngal,
                               by = "study_id",
                               all.x = TRUE)

# creates a new variable of the ang-2 concentration normalized to urine creatinine (pg/mg)
 clinical_lab_msd <- clinical_lab_msd %>%
   mutate(ucr_norm_ang2 = (ang2_concentration/(ucr_mg_dL*(1/100)))) %>% 
   mutate(ucr_norm_kim1 = (ukim1_pg_mL/(ucr_mg_dL*(1/100)))) %>%
   mutate(ucr_norm_ngal = (ungal_concentration/(ucr_mg_dL*(1/100))))
```

```{r}
# # creates a histogram of the normalized ang-2 concentrations
# clinical_lab_msd_no_outlier <- clinical_lab_msd %>%
#   filter(study_id != 278)
#   
# histogram <- ggplot(data = clinical_lab_msd,
#                     mapping = aes(x = ucr_norm_ang2)) +
#   geom_histogram()
# histogram
# 
# histogram <- ggplot(data = clinical_lab_msd_no_outlier,
#                     mapping = aes(x = ucr_norm_ang2)) +
#   geom_histogram()
# histogram
# 
# clinical_lab_msd$aki_hosp <- as.character(clinical_lab_msd$aki_hosp)
# # boxplot of normalized ang-2 concentrations and AKI within 7 days
# ang_hosp_aki <- ggplot(data = clinical_lab_msd,
#                      mapping = aes(x = aki_hosp,
#                                    y = log2(ucr_norm_ang2))) + 
#   geom_boxplot()
# ang_hosp_aki
# 
# clinical_lab_msd$dialysis_study <- as.character(clinical_lab_msd$dialysis_study)
# # boxplot of normalized ang-2 conc and dialysis within 7 days
# ang_7d_dialysis <- ggplot(data = clinical_lab_msd,
#                           mapping = aes(x = dialysis_study,
#                                         y = log2(ucr_norm_ang2))) +
#   geom_boxplot()
# ang_7d_dialysis
```

```{r, include = TRUE}
# # Correlation plot of ucr_norm_ang2 to Kim1 for Zoie to look at
# norm_ang2_kim1 <- ggplot(data = clinical_lab_msd,
#                          mapping = aes(
#                            y = log2(ucr_norm_kim1),
#                            x = log2(ucr_norm_ang2)))
# 
# norm_ang2_kim1 + 
#   geom_smooth(method = "lm",
#               se = FALSE) + 
#   geom_point() +
#   stat_regline_equation() +
#   stat_cor(label.y = 15,
#            aes(label = ..r.label..)) +
#   theme_classic() +
#   labs(x = "Log2(Normalized uAng-2)",
#        y = "Log2(Normalized uKim-1)")
```
Looking into the data by tertiles based on Ang-2 normalized to urine creatinine levels
```{r}
# splitting data into tertiles on ucr_norm_ang2
clinical_lab_msd$ucr_norm_ang2_tertile <- as.numeric(cut_number(clinical_lab_msd$ucr_norm_ang2,
                                        3))
clinical_lab_msd %>% count(ucr_norm_ang2_tertile)
# clinical_lab_msd$ucr_norm_ang2_tertile <- as.factor(clinical_lab_msd$ucr_norm_ang2_tertile)
# clinical_lab_msd$aki_hosp <- as.factor(clinical_lab_msd$aki_hosp)
```

```{r, include = TRUE}
# # Looks at the 7 day aki split by normalized ang-2 tertiles
# aki_norm_ang2 <- ggplot(data = clinical_lab_msd,
#                         mapping = aes(fill = ucr_norm_ang2_tertile,
#                                       x = aki_hosp))
# aki_norm_ang2 + 
#   geom_bar(position = "dodge")
# 
# # Looks at the 7 day dialysis split by normalized ang-2 tertiles
# dialysis_norm_ang2 <- ggplot(data = clinical_lab_msd,
#                              mapping = aes(fill = ucr_norm_ang2_tertile,
#                                            x = dialysis_study))
# dialysis_norm_ang2 +
#   geom_bar(position = "dodge")
# 
# boxes_norm_ang2 <- ggplot(data = clinical_lab_msd,
#                           mapping = aes(x = ucr_norm_ang2_tertile,
#                                         y = log2(ucr_norm_ang2)))
# boxes_norm_ang2 + geom_boxplot()
```

```{r, include = TRUE}
# # 10/5/2023
# # generates a scatterplot of ngal vs normalized ang2
# ngal_uang2 <- ggplot(data = clinical_lab_msd,
#                      mapping = aes(x = log2(ucr_norm_ang2),
#                                    y = log2(ucr_norm_ngal)))
# 
# ngal_uang2  + 
#   geom_smooth(method = "lm",
#               se = FALSE) + 
#   geom_point() +
#   stat_regline_equation(label.y = 23) +
#   stat_cor(label.y = 22,
#            aes(label = ..r.label..)) +
#   theme_classic() +
#   labs(x = "Log2(Normalized uAng-2)",
#        y = "Log2(Normalized uNGAL)")
```

```{r, include = TRUE}
# pang2 <- read_excel("B:/CHROME/Plasma/Ang1Ang2/2023-02-08/CHR_Ang1Ang2_20230209_ALL_PLATES_ANALYZED.xlsx",
#                     sheet = "WideSampleData")
# colnames(pang2) <- c("study_id",
#                      "PE_ang1_v1_concentration",
#                      "PE_ang2_v1_concentration")
# 
# clinical_lab_msd <- left_join(clinical_lab_msd,
#                                pang2,
#                                by = "study_id",
#                                all.x = TRUE)
# 
# clinical_lab_msd$PE_ang1_v1_concentration <- as.numeric(clinical_lab_msd$PE_ang1_v1_concentration)
# clinical_lab_msd$PE_ang2_v1_concentration <- as.numeric(clinical_lab_msd$PE_ang2_v1_concentration)
# 
# 
# # generates a scatterplot of ngal vs normalized ang2
# peang2_uang2 <- ggplot(data = clinical_lab_msd,
#                      mapping = aes(x = log2(ucr_norm_ang2),
#                                    y = log2(PE_ang2_v1_concentration)))
# 
# peang2_uang2  + 
#   geom_smooth(method = "lm",
#               se = FALSE) + 
#   geom_point() +
#   stat_regline_equation(label.y = 15.5) +
#   stat_cor(label.y = 15,
#            aes(label = ..r.label..)) +
#   theme_classic() +
#   labs(x = "Log2(Normalized uAng-2)",
#        y = "Log2(Plasma Ang-2)")
```

```{r, include = TRUE, echo = TRUE}
# # finds overlap between ang2 and normalized ang 2 tertiles
# ang2_tertile <- clinical_lab_msd %>%
#   dplyr::select(c("study_id",
#                   "ang2_tertile")) %>%
#   filter(ang2_tertile == 3)
# 
# norm_ang2_tertile <- clinical_lab_msd %>%
#   dplyr::select(c("study_id",
#                   "ucr_norm_ang2_tertile")) %>%
#   filter(ucr_norm_ang2_tertile == 3)
# 
# tertile_overlap <- inner_join(ang2_tertile,
#                               norm_ang2_tertile,
#                               by = "study_id")
# 
# length(tertile_overlap$study_id)
```
<br>
We see very high overlap between the highest Ang-2 Concentration tertiles and the highest normalized Ang-2 tertiles. 
53/64 patients are in the highest tertiles for both.
<br>

```{r}
# table 1 data finding
colnames(clinical_lab_msd)

clinical_lab_msd %>%
  group_by(ucr_norm_ang2_tertile) %>%
  summarise_at(vars(age),
               list(mean = mean,
                    sd = sd))

clinical_lab_msd %>%
  drop_na(sofa_d0) %>%
  group_by(ucr_norm_ang2_tertile) %>%
  summarise_at(vars(sofa_d0),
               list(mean = mean,
                    sd = sd))

clinical_lab_msd %>%
  drop_na(apacheiii_score) %>%
  group_by(ucr_norm_ang2_tertile) %>%
  summarise_at(vars(apacheiii_score),
               list(mean = mean,
                    sd = sd))

clinical_lab_msd %>%
  drop_na(bmi) %>%
  group_by(ucr_norm_ang2_tertile) %>%
  summarise_at(vars(bmi),
               list(mean = mean,
                    sd = sd))

# finds variable breakdown in each tertile
tertile <- c(1,2,3)

counter <- function(tertile) {
  clinical_lab_msd %>%
    filter(ucr_norm_ang2_tertile == tertile) %>%
    count()
}

lapply(tertile,
       counter)
```

```{r}
# # odds ratio
# # splits out data into individual tertiles
# or_lt <- clinical_lab_msd %>%
#   filter(ucr_norm_ang2_tertile == 1)
# or_mt <- clinical_lab_msd %>%
#   filter(ucr_norm_ang2_tertile == 2)
# or_ht <- clinical_lab_msd %>%
#   filter(ucr_norm_ang2_tertile == 3)
# 
# # new variable establishing lowest tertile as a reference 
# or_lt$tertile_or <- 0
# or_mt$tertile_or <- 1
# or_ht$tertile_or <- 1
# 
# # binds together the tertiles for analysis
# or_mt_lt <- rbind(or_lt,
#                  or_mt)
# or_ht_lt <- rbind(or_lt,
#                   or_ht)
# or_mt_ht <- rbind(or_mt,
#                   or_ht)
# 
# # runs a binomial regression with several variables for covariate adjustment 
# result <- glm(aki_hosp ~ tertile_or + age + male, 
#               data = or_mt_lt,
#               family = "binomial")
# 
# summary(result)
# exp(cbind(coef(result), confint.default(result)))
# 
# result <- glm(aki_hosp ~ tertile_or + age + male, 
#               data = or_ht_lt,
#               family = "binomial")
# 
# summary(result) 
# exp(cbind(coef(result), confint.default(result)))
```

```{r}
# relative risk
# converts several variables from factor to numeric
clinical_lab_msd$aki_hosp <- as.numeric(clinical_lab_msd$aki_hosp)
clinical_lab_msd$dialysis_study <- as.numeric(clinical_lab_msd$dialysis_study)

# # creates a new column that is a combination of dialysis and mortality
# clinical_lab_msd <- clinical_lab_msd %>%
#   mutate(death_dialysis = (dialysis_study + death_hospital))
# 
# # subtracts 1 from patients who died and had dialysis to create the larger group
# clinical_lab_msd$death_dialysis[clinical_lab_msd$death_dialysis == 2] <- 1
```


```{r}
# Ang-2 Relative Risk

# splits out data into individual tertiles
rr_lt <- clinical_lab_msd %>%
  filter(ucr_norm_ang2_tertile == 1)
# summary(rr_lt$ang2_concentration)

rr_mt <- clinical_lab_msd %>%
  filter(ucr_norm_ang2_tertile == 2)
# summary(rr_mt$ang2_concentration)

rr_ht <- clinical_lab_msd %>%
  filter(ucr_norm_ang2_tertile == 3)
# summary(rr_ht$ang2_concentration)

# new variable establishing lowest tertile as a reference for relative risk
rr_lt$tertile_rr <- 0
rr_mt$tertile_rr <- 1
rr_ht$tertile_rr <- 1

# binds together the tertiles for analysis
rr_mt_lt <- rbind(rr_lt,
                 rr_mt)
rr_ht_lt <- rbind(rr_lt,
                  rr_ht)
rr_mt_ht <- rbind(rr_mt,
                  rr_ht)

# runs a poisson regression with covariate adjustment to find relative risk
# aki_hosp
# unadjusted model
result <- glm(aki_hosp ~ tertile_rr, 
              data = rr_mt_lt,
              family = "poisson")

model_parameters(result,
                 exponentiate = TRUE,
                 vcov = "HC",
                 vcov_args = list(type = "HC0"))

# regression on lowest/middle tertile with adjustment
result <- glm(aki_hosp ~ tertile_rr + age + male + bmi + mv + covid_confirmed + myrace,
              data = rr_mt_lt,
              family = "poisson")

model_parameters(result,
                 exponentiate = TRUE,
                 vcov = "HC",
                 vcov_args = list(type = "HC0"))

# unadjusted 
result <- glm(aki_hosp ~ tertile_rr, 
              data = rr_ht_lt,
              family = "poisson")

model_parameters(result,
                 exponentiate = TRUE,
                 vcov = "HC",
                 vcov_args = list(type = "HC0"))

# regression on lowest/highest tertile with adjustment 
result <- glm(aki_hosp ~ tertile_rr + age + male + bmi + mv + covid_confirmed + myrace, 
              data = rr_ht_lt,
              family = "poisson")

model_parameters(result,
                 exponentiate = TRUE,
                 vcov = "HC",
                 vcov_args = list(type = "HC0"))

# dialysis_study
# # regression on lowest/middle tertile without adjustment
result <- glm(dialysis_study ~ tertile_rr, 
              data = rr_mt_lt,
              family = "poisson")

model_parameters(result,
                 exponentiate = TRUE,
                 vcov = "HC",
                 vcov_args = list(type = "HC0"))

# regression on lowest/middle tertile with adjustment
result <- glm(dialysis_study ~ tertile_rr + age + male + bmi + mv + covid_confirmed + myrace, 
              data = rr_mt_lt,
              family = "poisson")

model_parameters(result,
                 exponentiate = TRUE,
                 vcov = "HC",
                 vcov_args = list(type = "HC0"))

# regression on lowest/highest tertile without adjustment
result <- glm(dialysis_study ~ tertile_rr, 
              data = rr_ht_lt,
              family = "poisson")

model_parameters(result,
                 exponentiate = TRUE,
                 vcov = "HC",
                 vcov_args = list(type = "HC0"))

# regression on lowest/highest tertile with adjustment
result <- glm(dialysis_study ~ tertile_rr + age + male + bmi + mv + covid_confirmed + myrace, 
              data = rr_ht_lt,
              family = "poisson")

model_parameters(result,
                 exponentiate = TRUE,
                 vcov = "HC",
                 vcov_args = list(type = "HC0"))

# death_hospital
# # regression on lowest/middle tertile without adjustment
result <- glm(death_hospital ~ tertile_rr, 
              data = rr_mt_lt,
              family = "poisson")

model_parameters(result,
                 exponentiate = TRUE,
                 vcov = "HC",
                 vcov_args = list(type = "HC0"))

# regression on lowest/middle tertile with adjustment
result <- glm(death_hospital ~ tertile_rr + age + male + bmi + mv + covid_confirmed + myrace, 
              data = rr_mt_lt,
              family = "poisson")

model_parameters(result,
                 exponentiate = TRUE,
                 vcov = "HC",
                 vcov_args = list(type = "HC0"))

# regression on lowest/highest tertile without adjustment
result <- glm(death_hospital ~ tertile_rr, 
              data = rr_ht_lt,
              family = "poisson")

model_parameters(result,
                 exponentiate = TRUE,
                 vcov = "HC",
                 vcov_args = list(type = "HC0"))

# regression on lowest/highest tertile with adjustment
result <- glm(death_hospital ~ tertile_rr + age + male + bmi + mv + covid_confirmed + myrace, 
              data = rr_ht_lt,
              family = "poisson")

model_parameters(result,
                 exponentiate = TRUE,
                 vcov = "HC",
                 vcov_args = list(type = "HC0"))
```


```{r}
# # rr by SD
# # filters out the high outlier
# clinical_lab_msd_no_outlier <- clinical_lab_msd %>%
#   filter(study_id != 278)
# 
# # standard scales the data
# clinical_lab_msd_no_outlier$ucr_norm_ang2_scaled <- scale(clinical_lab_msd_no_outlier$ucr_norm_ang2)
# clinical_lab_msd$ucr_norm_ang2_scaled <- scale(clinical_lab_msd$ucr_norm_ang2)
# 
# # Hospital AKI
# # unadjusted model
# result <- glm(aki_hosp ~ ucr_norm_ang2_scaled, 
#               data = clinical_lab_msd_no_outlier,
#               family = "poisson")
# 
# model_parameters(result,
#                  exponentiate = TRUE,
#                  vcov = "HC",
#                  vcov_args = list(type = "HC0"))
# 
# result <- glm(aki_hosp ~ ucr_norm_ang2_scaled, 
#               data = clinical_lab_msd,
#               family = "poisson")
# 
# model_parameters(result,
#                  exponentiate = TRUE,
#                  vcov = "HC",
#                  vcov_args = list(type = "HC0"))
# 
# # adjusted model
# result <- glm(aki_hosp ~ ucr_norm_ang2_scaled + age + male + bmi + mv + covid_confirmed + myrace, 
#               data = clinical_lab_msd_no_outlier,
#               family = "poisson")
# 
# model_parameters(result,
#                  exponentiate = TRUE,
#                  vcov = "HC",
#                  vcov_args = list(type = "HC0"))
# 
# result <- glm(aki_hosp ~ ucr_norm_ang2_scaled + age + male + bmi + mv + covid_confirmed + myrace, 
#               data = clinical_lab_msd,
#               family = "poisson")
# 
# model_parameters(result,
#                  exponentiate = TRUE,
#                  vcov = "HC",
#                  vcov_args = list(type = "HC0"))
```

```{r, include = TRUE}
# # creates a histogram of the log2 transformed indexed uAng-2 data
# historgram <- ggplot(data = clinical_lab_msd,
#                      mapping = aes(x = log2(ucr_norm_ang2)))
# 
# historgram + geom_histogram() +
#   labs(x = "Log2(Indexed uAng-2)")
```

```{r, include = TRUE}
# # grubbs test for outlier
# library(outliers)
# 
# grubbs.test(clinical_lab_msd$ucr_norm_ang2, 
#             two.sided = FALSE)
```
CHR_287 is an outlier according to the grubbs tes (p-value < 2.2 e-16)

```{r}
# NGAL Relative Risk

# splitting data in to ngal tertiles
clinical_lab_msd$ucr_norm_ngal_tertile <- as.numeric(cut_number(clinical_lab_msd$ucr_norm_ngal,
                                        3))
clinical_lab_msd_ngal <- clinical_lab_msd %>%
  drop_na(ucr_norm_ngal_tertile)
clinical_lab_msd_ngal %>% count(ucr_norm_ngal_tertile)

# sets the data for a relative risk analysis
# splits out data into individual tertiles
rr_lt <- clinical_lab_msd_ngal %>%
  filter(ucr_norm_ngal_tertile == 1)

rr_mt <- clinical_lab_msd_ngal %>%
  filter(ucr_norm_ngal_tertile == 2)

rr_ht <- clinical_lab_msd_ngal %>%
  filter(ucr_norm_ngal_tertile == 3)

# new variable establishing lowest tertile as a reference for relative risk
rr_lt$tertile_rr <- 0
rr_mt$tertile_rr <- 1
rr_ht$tertile_rr <- 1

# binds together the tertiles for analysis
rr_mt_lt <- rbind(rr_lt,
                 rr_mt)
rr_ht_lt <- rbind(rr_lt,
                  rr_ht)
rr_mt_ht <- rbind(rr_mt,
                  rr_ht)

# regressions

# aki_hosp
# unadjusted model
result <- glm(aki_hosp ~ tertile_rr, 
              data = rr_mt_lt,
              family = "poisson")

model_parameters(result,
                 exponentiate = TRUE,
                 vcov = "HC",
                 vcov_args = list(type = "HC0"))

# regression on lowest/middle tertile with adjustment
result <- glm(aki_hosp ~ tertile_rr + age + male + bmi + mv + covid_confirmed + myrace,
              data = rr_mt_lt,
              family = "poisson")

model_parameters(result,
                 exponentiate = TRUE,
                 vcov = "HC",
                 vcov_args = list(type = "HC0"))

# unadjusted 
result <- glm(aki_hosp ~ tertile_rr, 
              data = rr_ht_lt,
              family = "poisson")

model_parameters(result,
                 exponentiate = TRUE,
                 vcov = "HC",
                 vcov_args = list(type = "HC0"))

# regression on lowest/highest tertile with adjustment 
result <- glm(aki_hosp ~ tertile_rr + age + male + bmi + mv + covid_confirmed + myrace, 
              data = rr_ht_lt,
              family = "poisson")

model_parameters(result,
                 exponentiate = TRUE,
                 vcov = "HC",
                 vcov_args = list(type = "HC0"))

# dialysis_study
# # regression on lowest/middle tertile without adjustment
result <- glm(dialysis_study ~ tertile_rr, 
              data = rr_mt_lt,
              family = "poisson")

model_parameters(result,
                 exponentiate = TRUE,
                 vcov = "HC",
                 vcov_args = list(type = "HC0"))

# regression on lowest/middle tertile with adjustment
result <- glm(dialysis_study ~ tertile_rr + age + male + bmi + mv + covid_confirmed + myrace, 
              data = rr_mt_lt,
              family = "poisson")

model_parameters(result,
                 exponentiate = TRUE,
                 vcov = "HC",
                 vcov_args = list(type = "HC0"))

# regression on lowest/highest tertile without adjustment
result <- glm(dialysis_study ~ tertile_rr, 
              data = rr_ht_lt,
              family = "poisson")

model_parameters(result,
                 exponentiate = TRUE,
                 vcov = "HC",
                 vcov_args = list(type = "HC0"))

# regression on lowest/highest tertile with adjustment
result <- glm(dialysis_study ~ tertile_rr + age + male + bmi + mv + covid_confirmed + myrace, 
              data = rr_ht_lt,
              family = "poisson")

model_parameters(result,
                 exponentiate = TRUE,
                 vcov = "HC",
                 vcov_args = list(type = "HC0"))

# death_hospital
# # regression on lowest/middle tertile without adjustment
result <- glm(death_hospital ~ tertile_rr, 
              data = rr_mt_lt,
              family = "poisson")

model_parameters(result,
                 exponentiate = TRUE,
                 vcov = "HC",
                 vcov_args = list(type = "HC0"))

# regression on lowest/middle tertile with adjustment
result <- glm(death_hospital ~ tertile_rr + age + male + bmi + mv + covid_confirmed + myrace, 
              data = rr_mt_lt,
              family = "poisson")

model_parameters(result,
                 exponentiate = TRUE,
                 vcov = "HC",
                 vcov_args = list(type = "HC0"))

# regression on lowest/highest tertile without adjustment
result <- glm(death_hospital ~ tertile_rr, 
              data = rr_ht_lt,
              family = "poisson")

model_parameters(result,
                 exponentiate = TRUE,
                 vcov = "HC",
                 vcov_args = list(type = "HC0"))

# regression on lowest/highest tertile with adjustment
result <- glm(death_hospital ~ tertile_rr + age + male + bmi + mv + covid_confirmed + myrace, 
              data = rr_ht_lt,
              family = "poisson")

model_parameters(result,
                 exponentiate = TRUE,
                 vcov = "HC",
                 vcov_args = list(type = "HC0"))

# finds variable breakdown in each tertile
tertile <- c(1,2,3)

counter <- function(tertile) {
   clinical_lab_msd_ngal %>%
   filter(ucr_norm_ngal_tertile == tertile) %>%
   count(death_hospital)
}

lapply(tertile,
       counter)
```

```{r}
# KIM-1 Relative Risk

# splitting data in to kim1 tertiles
clinical_lab_msd$ucr_norm_kim1_tertile <- as.numeric(cut_number(clinical_lab_msd$ucr_norm_kim1,
                                        3))
clinical_lab_msd %>% count(ucr_norm_kim1_tertile)

# sets the data for a relative risk analysis
# splits out data into individual tertiles
rr_lt <- clinical_lab_msd %>%
  filter(ucr_norm_kim1_tertile == 1)

rr_mt <- clinical_lab_msd %>%
  filter(ucr_norm_kim1_tertile == 2)

rr_ht <- clinical_lab_msd %>%
  filter(ucr_norm_kim1_tertile == 3)

# new variable establishing lowest tertile as a reference for relative risk
rr_lt$tertile_rr <- 0
rr_mt$tertile_rr <- 1
rr_ht$tertile_rr <- 1

# binds together the tertiles for analysis
rr_mt_lt <- rbind(rr_lt,
                 rr_mt)
rr_ht_lt <- rbind(rr_lt,
                  rr_ht)
rr_mt_ht <- rbind(rr_mt,
                  rr_ht)

# regressions

# aki_hosp
# unadjusted model
result <- glm(aki_hosp ~ tertile_rr, 
              data = rr_mt_lt,
              family = "poisson")

model_parameters(result,
                 exponentiate = TRUE,
                 vcov = "HC",
                 vcov_args = list(type = "HC0"))

# regression on lowest/middle tertile with adjustment
result <- glm(aki_hosp ~ tertile_rr + age + male + bmi + mv + covid_confirmed + myrace,
              data = rr_mt_lt,
              family = "poisson")

model_parameters(result,
                 exponentiate = TRUE,
                 vcov = "HC",
                 vcov_args = list(type = "HC0"))

# unadjusted 
result <- glm(aki_hosp ~ tertile_rr, 
              data = rr_ht_lt,
              family = "poisson")

model_parameters(result,
                 exponentiate = TRUE,
                 vcov = "HC",
                 vcov_args = list(type = "HC0"))

# regression on lowest/highest tertile with adjustment 
result <- glm(aki_hosp ~ tertile_rr + age + male + bmi + mv + covid_confirmed + myrace, 
              data = rr_ht_lt,
              family = "poisson")

model_parameters(result,
                 exponentiate = TRUE,
                 vcov = "HC",
                 vcov_args = list(type = "HC0"))

# dialysis_study
# # regression on lowest/middle tertile without adjustment
result <- glm(dialysis_study ~ tertile_rr, 
              data = rr_mt_lt,
              family = "poisson")

model_parameters(result,
                 exponentiate = TRUE,
                 vcov = "HC",
                 vcov_args = list(type = "HC0"))

# regression on lowest/middle tertile with adjustment
result <- glm(dialysis_study ~ tertile_rr + age + male + bmi + mv + covid_confirmed + myrace, 
              data = rr_mt_lt,
              family = "poisson")

model_parameters(result,
                 exponentiate = TRUE,
                 vcov = "HC",
                 vcov_args = list(type = "HC0"))

# regression on lowest/highest tertile without adjustment
result <- glm(dialysis_study ~ tertile_rr, 
              data = rr_ht_lt,
              family = "poisson")

model_parameters(result,
                 exponentiate = TRUE,
                 vcov = "HC",
                 vcov_args = list(type = "HC0"))

# regression on lowest/highest tertile with adjustment
result <- glm(dialysis_study ~ tertile_rr + age + male + bmi + mv + covid_confirmed + myrace, 
              data = rr_ht_lt,
              family = "poisson")

model_parameters(result,
                 exponentiate = TRUE,
                 vcov = "HC",
                 vcov_args = list(type = "HC0"))

# death_hospital
# # regression on lowest/middle tertile without adjustment
result <- glm(death_hospital ~ tertile_rr, 
              data = rr_mt_lt,
              family = "poisson")

model_parameters(result,
                 exponentiate = TRUE,
                 vcov = "HC",
                 vcov_args = list(type = "HC0"))

# regression on lowest/middle tertile with adjustment
result <- glm(death_hospital ~ tertile_rr + age + male + bmi + mv + covid_confirmed + myrace, 
              data = rr_mt_lt,
              family = "poisson")

model_parameters(result,
                 exponentiate = TRUE,
                 vcov = "HC",
                 vcov_args = list(type = "HC0"))

# regression on lowest/highest tertile without adjustment
result <- glm(death_hospital ~ tertile_rr, 
              data = rr_ht_lt,
              family = "poisson")

model_parameters(result,
                 exponentiate = TRUE,
                 vcov = "HC",
                 vcov_args = list(type = "HC0"))

# regression on lowest/highest tertile with adjustment
result <- glm(death_hospital ~ tertile_rr + age + male + bmi + mv + covid_confirmed + myrace, 
              data = rr_ht_lt,
              family = "poisson")

model_parameters(result,
                 exponentiate = TRUE,
                 vcov = "HC",
                 vcov_args = list(type = "HC0"))

counter <- function(tertile) {
   clinical_lab_msd_ngal %>%
   filter(ucr_norm_kim1_tertile == tertile) %>%
   count(aki_hosp)
}

lapply(tertile,
       counter)
```

